---
title: 'STAT 521: Assignment 8'
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacs <- c("kableExtra", "tidyverse")
sapply(pacs, require, character.only = TRUE)
```
**Make sure to show your computation and/or attach appropriate output.**

### Problem 1 (Computer exercise)
A person’s muscle mass is expected to decrease with age. To explore this relationship in women, a nutritionist randomly selected 15 women from each 10-year age group, beginning with age 40 and ending with age 79, and measured their muscle mass.

a)	Create a scatter plot between age and muscle mass. Choose variables for x- and y-axes appropriately. Is there any relationship between age and muscle mass? If so, is the association positive or negative? Does the relationship appear to be linear? Any outliers? **Describe your findings**.

b)	Obtain the Pearson correlation coefficient between age and muscle mass. Is the correlation significantly different from zero? Report the 95% confidence interval for $\rho$

c) Run simple regression. **Again, make sure to choose an appropriate variable for each of $X$ and $Y$**. Report the estimated regression equation. Attach the ANOVA table. What is the value of $R^2$ and its interpretation? What is a point estimate for $\sigma^2$? Produce a scatter plot with the fitted regression line.

d) What is a point estimate of the difference in the mean muscle mass for women differing in age by one year? Report its 95% confidence interval too.

e) Suppose you wish to predict muscle mass of a woman aged 60 based on your regression model. Calculate her predicted muscle mass. Report its 95% prediction interval too. If the woman has muscle mass of 105, what is the value of the residual for her?

f) Is it appropriate to estimate muscle mass of a woman aged 20 using the regression equation you obtained above? Discuss.

g) Produce a residual plot against fitted (predicted) values, as well as a normal probability plot of residuals. Are there any outliers? Are residuals normally distributed? Is there any non-linear pattern in residuals? How about the equal variance assumption?

### Problem 2
Complete the ANOVA table below for simple linear regression of $n = 27$ and answer the following questions.

\renewcommand{\arraystretch}{2}
```{r, echo = FALSE}
anova_tab <- data.frame(Source = c("Model", "Error", "Total"), df = rep("", 3), 
                        SS = rep("", 3), MS = c("840", "", ""), F = c("10.5", "", ""))
anova_tab %>%   
  kbl(align = "c") %>%
  column_spec(1, bold = TRUE) %>%
  column_spec(2, width = "8em") %>%
  column_spec(3, width = "8em") %>%
  column_spec(4, width = "8em", background = c("white", "white", "gray")) %>%
  column_spec(5, width = "8em", background = c("white", "gray", "gray")) %>%
  row_spec(0, bold = TRUE, italic = TRUE) %>%
  kable_paper("hover",full_width = F, latex_options = "HOLD_position")
```

a) Calculate $R^2$.

b) It is known that $X$ and $Y$ used in here are negatively associated. Using the information above, what is the correlation coefficient between $X$ and $Y$?

c) Means and standard deviations of $X$ and $Y$ are given below. Using this and part (b), obtain the estimated regression equation.

\renewcommand{\arraystretch}{1}
```{r, echo = FALSE}
mean_sd_tab <- data.frame(Mean = c(110.2, 55.0), SD = c(20.5, 8.2))
row.names(mean_sd_tab) <- c("X", "Y")

mean_sd_tab %>%   
  kbl() %>%
  column_spec(1, bold = TRUE) %>%
  column_spec(2, width = "3em") %>%
  column_spec(3, width = "3em") %>%
  row_spec(0, bold = TRUE) %>%
  kable_paper("hover",full_width = F, latex_options = "HOLD_position")
```


### Problem 3 (Biostats students only)

*Expected value of a function of random variables*:

In Assignment #5, you learned the joint probability density (or mass) function. Suppose random variables $X$ and $Y$ have a joint probability density/mass function, $f_{X, Y}(x, y)$. How can we calculate the expected value of any function of $X$ and $Y$, e.g., $E\left[ g\left(X, Y\right) \right]$?

**Definition**:

For discrete random variable $X$ and $Y$:

$$E\left[ g\left(X, Y\right) \right] = \sum_x\sum_y g(x, y) \, f_{X, Y}(x, y)$$

For continuous random variable $X$ and $Y$:

$$E\left[ g\left(X, Y\right) \right] = \int ^{\infty}_{-\infty}\int ^{\infty}_{-\infty} g(x, y) \, f_{X, Y}(x, y) \, dxdy$$

**Example**: Let $X$ and $Y$, both continuous have the joint probability density function:

$$ 
f_{X, Y}(x, y) =  
\begin{cases}
\frac{1}{6}(x + 4y), & 0 \le x \le 2, \; 0 \le y \le 1 \\
0, & \text{elsewhere}
\end{cases}
$$

We want to find $E(XY)$.

$$
\begin{aligned}
E(XY) &= \int ^{\infty}_{-\infty}\int ^{\infty}_{-\infty} xyf_{X, Y}(x, y) \, dxdy \\
      &= \int ^1_0 \int ^2_0 xy\frac{1}{6}(x + 4y) \, dxdy \\
      &= \int ^1_0 \int ^2_0 \frac{1}{6}(x^2y + 4xy^2) \, dxdy \\
      &= \int ^1_0 \left( \frac{1}{18}x^3y + \frac{1}{3}x^2y^2 \bigg] ^2_0  \right)dy \\
      &= \int ^1_0 \left( \frac{4}{9}y + \frac{4}{3}y^2 \right)dy =  \frac{2}{9}y^2 + \frac{4}{9}y^3 \bigg]^1_0 = \frac{2}{3}
\end{aligned}
$$
\bigskip

**Problem**: Let $X$ and $Y$ have the joint probability density function given by:

$$
 f_{X, Y}(x, y) =  
\begin{cases}
4xy, & 0 \le x \le 1, \; 0 \le y \le 1 \\
0, & \text{elsewhere}
\end{cases}
$$

a) Find $E(X)$ and $E(Y)$.

b) The covariance between X and Y can be written as:
$$cov(X, Y) = E \left[ (X- \mu_X)(Y- \mu_Y)\right] = E(XY) - \mu_X\mu_Y$$
\setlength{\leftskip}{2em}
where $\mu_X = E(X)$ and $\mu_Y = E(Y)$. Find $E(XY)$ and $cov(X, Y)$.
\setlength{\leftskip}{0em}

c) Use the part (b) to find the correlation coefficient between $X$ and $Y$: $\rho_{X, Y}$ Remember, from Lecture 10, the correlation coefficient is defined as:
$$\rho_{X, Y} = \frac{cov(X, Y)}{\sqrt{var(X)}\sqrt{var(Y)}}$$

\bigskip

### Problem 4 (Biostats students only)

*Marginal probability distributions, independence of RVs*

Suppose you have a joint probability distribution of two discrete random variables $X$ and $Y$, $p_{X, Y} (x, y)$, as follows (example taken from Assignment #5):

```{r echo = FALSE, results = 'asis'}
tab <- matrix(c(0.05, 0.56, 0.61, 0.10, 0.18, 0.28, 0.09, 0.02, 0.11, 0.24, 0.76, 1.00), ncol = 3, byrow = TRUE)
rownames(tab) <- c("0", "1", "2", "Total")
colnames(tab) <- c("0", "1", "Total")
tab %>% 
  kbl() %>%
  add_header_above(c("X" = 1, "Y" = 2, " " = 1)) %>% 
  column_spec(1, bold = T) %>%
  kable_paper("hover", full_width = F, latex_options = "HOLD_position")
```

How do we get the marginal probability distribution of $X$, that is, $p_X(x)$ ? This is easy
because:

$$
\begin{aligned}
p_X(0) &= Pr(X = 0) = p_{X, Y}(0, 0) + p_{X, Y}(0, 1) = 0.61 \\
p_X(1) &= Pr(X = 1) = p_{X, Y}(1, 0) + p_{X, Y}(1, 1) = 0.28 \\
p_X(2) &= Pr(X = 2) = p_{X, Y}(2, 0) + p_{X, Y}(2, 1) = 0.11
\end{aligned}
$$

In general, marginal probability mass functions (for discrete cases) are given by:

$$p_X(x) = \sum_y p_{X, Y}(x, y) \qquad \text{and} \qquad p_Y(y) = \sum_x p_{X, Y}(x, y)$$

Similarly, if $X$ and $Y$ are continuous, marginal density functions are:

$$f_X(x) = \int^{\infty}_{-\infty} f_{X, Y}(x, y)\,dy \qquad \text{and} \qquad f_Y(y) = \int^{\infty}_{-\infty} f_{X, Y}(x, y)\,dx$$
\bigskip

**Example**: Let $X$ and $Y$, both continuous, have the joint probability density function:

$$
f_{X, Y}(x, y) =  
\begin{cases}
\frac{1}{6}(x + 4y), & 0 \le x \le 2, \; 0 \le y \le 1 \\
0, & \text{elsewhere}
\end{cases}
$$

We want to find $f_X(x)$ and $f_Y(y)$.

For $0 \le x \le 2$,
$$
 f_X(x) = \int^{\infty}_{-\infty} f_{X, Y}(x, y)\,dy = \int^{1}_{0} \frac{1}{6}(x + 4y)\,dy = \frac{1}{6}xy + \frac{1}{3}y^2 \bigg] ^1_0 = \frac{x + 2}{6}
$$

and $f_X(x) = 0$ elsewhere. For $0 \le y \le 1$,

$$
 f_y(y) = \int^{\infty}_{-\infty} f_{X, Y}(x, y)\,dx = \int^{2}_{0} \frac{1}{6}(x + 4y)\,dx = \frac{1}{12}x^2 + \frac{2}{3}xy \bigg] ^2_0 = \frac{4y + 1}{3}
$$

and $f_Y(y) = 0$ elsewhere. 

\bigskip

**Problem**: Let $X$ and $Y$ have the joint probability density function given by:

$$
 f_{X, Y}(x, y) =  
\begin{cases}
4xy, & 0 \le x \le 1, \; 0 \le y \le 1 \\
0, & \text{elsewhere}
\end{cases}
$$

a) Find $f_X(x)$ and $f_Y(y)$.

b) If the joint density can be written as the product of individual density functions, i.e.,
$$f_{XY}(x, y) = f_X(x)f_Y(y)$$
\setlength{\leftskip}{2em}
then $X$ and $Y$ are said to be independent random variables. Are $X$ and $Y$ independent?
\setlength{\leftskip}{0em}
c) If random variables X and Y are independent, then we have this property:
$$E(XY) = E(X)E(Y)$$
\setlength{\leftskip}{2em}
Find $E(XY)$. Is your answer same as the one you get for Problem 3 part (b)?
\setlength{\leftskip}{0em}

\bigskip

**Additional notes**:
If random variables $X$ and $Y$ are independent, then the covariance between $X$ and $Y$ is zero and so is the correlation.

\begin{center}
If $X$ and $Y$ independent $\implies cov(X, Y) = 0$
\end{center}

But the converse is not necessarily true. You can have a correlation of zero between $X$ and $Y$ that are not independent. See the following joint probability distribution.

```{r echo = FALSE, results = 'asis'}
tab <- matrix(c("1/3", "0", "1/3", "2/3", "0", "1/3", "0", "1/3", "1/3", "1/3", "1/3", "1"), ncol = 4, byrow = TRUE)
rownames(tab) <- c("0", "1", "Total")
colnames(tab) <- c("0", "1", "2", "Total")
tab %>% 
  kbl(align = "r") %>%
  add_header_above(c("X" = 1, "Y" = 3, " " = 1)) %>% 
  column_spec(1, bold = T) %>%
  kable_paper("hover", full_width = F, latex_options = "HOLD_position")
```

Verify that $cov(X, Y) = E(XY) -E(X)E(Y) = \frac{1}{3} - \left( \frac{1}{3} \right) \left( 1 \right) = 0$. But $X$ and $Y$ are not independent because $Pr(X = 0 \,| \,Y = 0) \ne Pr(X = 0)$.

