---
title: 'STAT 521: Assignment 7'
output: 
  pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacs <- c("kableExtra", "tidyverse")
sapply(pacs, require, character.only = TRUE)
```
**Make sure to show your computation and/or attach appropriate output.**

### Problem 1

A researcher wishes to compare birth weights of infants among four groups of their mothers’ smoking status: non-smokers, ex-smokers, $<1/2$ pack/day, $\ge1/2$ pack/day. Data is given as follows (See Assignment7.sas/R/py for computer analysis):

\begin{center}
\begin{tabular}{ c | c | c | c | c | c}
 & Non-smokers & Ex-smokers & $< \frac{1}{2}$ pack/d & $\ge \frac{1}{2}$ pack/d & \\ 
 \hline
 & 8.56 & 7.39 & 5.97 & 7.03 & \\  
 & 8.47 & 8.64 & 6.77 & 5.24 & \\  
 & 6.39 & 8.54 & 7.26 & 6.14 & \\  
 & 9.26 & 5.37 & 5.74 & 6.74 & \\  
 & 7.98 & 9.21 & 8.74 & 6.62 & \\  
 & 6.84 & 6.63 & 6.30 & 7.37 & \\  
 & 7.87 &      & 5.52 & 4.94 & \\  
 &      &      &      & 6.34 & \\  
\hline
 $\sum y$ & 55.37 & 45.78 & 46.30 & 50.42 & $\sum\sum y = 197.87$ \\
\hline
 $\sum y^2$ & 444.00	& 359.81 & 313.68	& 322.75 & $\sum\sum y^2 = 1440.23$ \\
\hline
 $\bar y$ & 7.91   & 	7.63 &   6.61	&   6.30 & $ \bar{y}_{\cdot\cdot}= 7.07$ \\
\hline
 $n$      & 7      &   6    &   7    &   8    & 28 \\
\hline
\end{tabular}
\end{center}

a)	State the null and alternative hypotheses.

b)	Manually calculate sum of squares and complete the ANOVA table below.

\renewcommand{\arraystretch}{2}
```{r, echo = FALSE}
anova_tab <- data.frame(Source = c("Treatment", "Error", "Total"), df = rep("", 3), SS = rep("", 3), MS = rep("", 3), F = rep("", 3))
anova_tab %>%   
  kbl(align = "c") %>%
  column_spec(1, bold = TRUE) %>%
  column_spec(2, width = "8em") %>%
  column_spec(3, width = "8em") %>%
  column_spec(4, width = "8em", background = c("white", "white", "gray")) %>%
  column_spec(5, width = "8em", background = c("white", "gray", "gray")) %>%
  row_spec(0, bold = TRUE, italic = TRUE) %>%
  kable_paper("hover",full_width = F, latex_options = "HOLD_position")
```

c)	Test the hypothesis. Use $\alpha = 0.05$. **What is your conclusion?**

d)	What are the assumptions that you are making to conduct the above test?

e)	Use SAS, R or Python to verify your results in part (b) and (c). Run Tukey’s HSD test for multiple comparisons. Did you find any groups that were significantly different at $\alpha = 0.05$? For each of significant pairs (if any), report a 95% confidence interval of the mean difference. **Attach relevant output** and **report your findings**.

### Problem 2

A researcher conducted a one-way ANOVA to test the null hypothesis $H_0: \mu_1 = \mu_2 = \mu_3 = \mu_4$. The researcher had a balanced design with total number of subjects $N = 20$. Complete the ANOVA table below.

```{r, echo = FALSE}
anova_tab <- data.frame(Source = c("Treatment", "Error", "Total"), df = rep("", 3), 
                        SS = c("6750", "8000", ""), MS = rep("", 3), F = rep("", 3))
anova_tab %>%   
  kbl(align = "c") %>%
  column_spec(1, bold = TRUE) %>%
  column_spec(2, width = "8em") %>%
  column_spec(3, width = "8em") %>%
  column_spec(4, width = "8em", background = c("white", "white", "gray")) %>%
  column_spec(5, width = "8em", background = c("white", "gray", "gray")) %>%
  row_spec(0, bold = TRUE, italic = TRUE) %>%
  kable_paper("hover",full_width = F, latex_options = "HOLD_position")
```

### Problem 3 (Biostats students only)

*Uniform distribution*:

One of the simplest types of continuous distributions is the *uniform distribution*. The probability density function (PDF) of a uniform random variable $X$ is defined as:

$$f(x) = \begin{cases}
         \frac{1}{b - a}, & a < x < b \\
         0, & \text{elsewhere}
         \end{cases}
$$

```{r echo = FALSE, fig.height = 1.6, fig.width = 3, fig.align='center', fig.cap = "The probability density fucntion of $Uniform(a, b)$"}
par(mar = c(1, 1, .2, 1))
plot(c(4, 0), c(0, 2), type = "n", axes = FALSE, xlab = "", ylab = "")
abline(h = 0); abline(v = 0.3)
segments(0.3, 1, 1, 1, lty = 2)
segments(1, 1, 3, 1, col = "red", lwd = 2)
segments(1, 1, 1, 0, lty = 2)
segments(3, 1, 3, 0, lty = 2)
text(0.1, 1.9, expression(italic(f(x))), cex = 0.5, family = "serif")
text(0.05, 1, expression(italic(frac(1, b - a))), cex = 0.5, family = "serif")
mtext(expression(italic(a)), 1, at = 1, cex = 0.5, family = "serif")
mtext(expression(italic(b)), 1, at = 3, cex = 0.5, family = "serif")
mtext(expression(italic(x)), 1, at = 4, cex = 0.5, family = "serif")

```

A random variable $X$ can take any value between $a$ and $b$, and the density is constant over this range. We write this as:

$$X \sim Uniform(a, b)$$

A special case of the uniform distribution is when $a = 0$ and $b = 1$: $X \sim Uniform(0, 1)$. This is called the *standard uniform distribution*. The standard uniform distribution has a PDF of $f(x) = 1$ for $0 < x < 1$ and $f(x) = 0$ otherwise.

**Question**: Use the definition of the expected value $E(X) = \int^{\infty}_{-\infty} xf(x) \,dx$ (see Assignment #3) and show that $E(X) = \frac{a + b}{2}$ and $var(X) = \frac{(b - a) ^ 2}{12}$ if $X \sim Uniform(a, b)$.

### Problem 4 (Biostats students only)

*Monte Carlo integration using R*

Suppose we wish to solve a definite integral $I = \int^b_a g(x) \, dx$, $a < b$. Note that $g(x)$ is any one-dimensional function, not necessarily a PDF. Sometimes $g(x)$ is very complex and there may not be a closed-form solution. We can rewrite the integral as:

$$I = \int^b_a g(x) \, dx = I = \int^b_a \frac{g(x)}{f(x)} f(x) \, dx$$

If we can find a random variable that has a well-known PDF of $f(x)$ whose support is $a < x < b$, then by the definition of the expected value, the integral is equivalent to:

$$I = \int^b_a \frac{g(x)}{f(x)} f(x) \, dx = E\left[ \frac{g(x)}{f(x)} \right]$$

in which a random variable $X$ has a PDF $f(x) \ge 0$ for $a < x < b$ and $f(x) = 0$ otherwise. An obvious choice of $X$ is a uniform random variable $X \sim Uniform(a, b)$ with $f(x) = \frac{1}{b - a}$ for $a < x < b$. Then it follows:

$$I = E\left[ \frac{g(x)}{f(x)} \right] = E\left[ \frac{g(x)}{1 / (b - a)} \right] = (b - a) E \left[ g(x) \right]$$

What does this imply? Remember, from Assignment #6, you can approximate $E \left[ g(x) \right]$ as long as you can generate a large number of random values from $X \sim Uniform(a, b)$. If we take the average of all the random values of $g(x)$ and multiply by $(b - a)$, we can approximate the definite integral $I = \int^b_a g(x) \, dx$.

\bigskip

**Example**: Suppose we want to solve:

$$I = \int^2_0 \frac{2e^{-2x}}{\left(1 + e^{-2x}\right) ^ 2} \, dx$$ 

Let $g(x) = \frac{2e^{-2x}}{\left(1 + e^{-2x}\right) ^ 2}$ and $X$ be a uniform random variable $X \sim Uniform(0, 2)$ whose PDF is $f(x) = \frac{1}{2}$ for $0 < x < 2$. Then we can approximate the integral $I$ as $2E\left[ g(X) \right]$. 

Now let’s do this with R. First I’m going to produce one million random numbers from $X \sim Uniform(0, 2)$. This can be done using `runif()` function. Its syntax is `runif(n, a, b)` where `n` is the number of random values you want to generate, and `a` and `b` are the range of $x$. In the code below, I assign values $a = 0$ and $b = 2$.

```{r include=FALSE}
set.seed(123)

```
```{r}
# Generate 1 million random number from X ~ Unif(0, 2)
a <- 0
b <- 2
x <- runif(10 ^ 6, a, b)
head(x)
```

Then for each value of $x$, I calculate $g(x)$.  For example, for the first uniform random value of `r x[1]`, we have $g(`r x[1]`) = \frac{2e^{-2(`r x[1]`)}}{\left(1 + e^{-2(`r x[1]`)}\right) ^ 2} \approx 0.365$. For $e^x$, use `exp()` function.

```{r}
# Calculate g(x)
g <- 2 * exp(-2 * x) / (1 + exp(-2 * x)) ^ 2
head(g)
```

Now, to approximate $I = (b - a) E \left[ g(x) \right]$, we calculate the mean of `g` and then multiply by $b-a = 2$.

```{r}
# Take the average and multiply by (b - a)
(b - a) * mean(g)
```

In this case, I got $I \approx `r (b - a) * mean(g)`$. This is close to the correct answer of $I = \int^2_0 \frac{2e^{-2x}}{\left(1 + e^{-2x}\right) ^ 2} \, dx = 0.4820138$.

\bigskip

**Here's an exercise for you**: 

Approximate:

$$I = \int^{0.3}_{0.2} 15(3x)^4 \, e^{-(3x^5)} \, dx$$

Please submit your R code as well.

<br><br><br><br><br>
